Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-08-20 how.tex
Updates from ShareLaTeX

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, those improvements   reduce
 the expected values of defect counts and  runtimes to    
 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, case-based reasoning.
  \end{IEEEkeywords}

@@

This is an open and pressing issue.
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself; e.g.
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
To generate those plans, we propose the ``HOW'' case-based reasoner (CBR).
HOW's  ``plans'' are  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every test case $Z$, HOW proposes a  plan $\delta$ to
  adjust   attributes $Z_j$:
\[
Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, HOW might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).
HOW's  plans displace the test case into some region
of the training data. This is important since we will use that training data
to build a predictor to assess the impact of that plan.
To put that another way, when a novel test case is presented,
HOW says 
\begin{quote}{\em ``I may not know too much about that, but once you change into something like
what I've already seen, then I can suggest how to improve it''.}
\end{quote}
After generating the plans, the next question is ``how to assess them?''.  
Some organizations have the resources to 
run repeated trials to assess  different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
then check in 12 months time
 which teams have fewer defects than the others.

@@

This paper proposes another, less resource intensive,  method for assessing the effectiveness of proposed
changes to a project.  
We  use a prediction system (built by data mining) as a verification
tool for planners. In domains where  data miners can build
predictors of adequate performance, then those predictors can assess the value
of proposed changes.
This paper applies this proposal to show that   HOW's plans are quite effective and  can
 significantly reduce the  predicted
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
The rest of this paper presents the background to this work in Section~2; the design of HOW in Section~3;
then our experimental results  in Section~4.  Our conclusion is that we can plan for the future in
software engineering, provided the domain data supports adequate predictors.

@@

However, not all domains come with ready-to-use models. For example, 
consider software defect prediction and all the intricate issues that may lead defects in a product. A model that includes {\em all} those
potential issues would be very large and complex. Further,
the empirical data required to validate any/all parts
of that model can be hard to find.

@@

In summary, for domains with readily accessible models, we recommend
the kinds of tools that are widely used in the search-based
software engineering community such as GALE,NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like HOW.

\subsection{Planning and AI}
HOW is a {\em probabilistic} planner, which is a different approach to that taken by
classical AI. 
That kind of planning is a logical procedure~\cite{Fikes1971}
that seeks an ordering on {\em operators} to take some domain
{\em state} from a {\em start} state to a {\em  goal} state.
This classical logical approach is known to suffer from
computational bottlenecks~\cite{Bylander1994}. Hence,
like this paper,
many researchers take a probabilistic approach.
HOW's commitment to minimal displacement (and the fully-automated
modeling approach) of this paper makes it stand apart from, say,
  the Bayesian/Markov chain approach to planning of Dean et al.~\cite{dean1995planning}.

\subsection{Case-based Reasoning}
Case-based reasoning (CBR) has been extensively
explored in the AI literature
(where it has also bee called
 reasoning via analogy or instance-based reasoning).
CBR seeks to emulate human recollection and adaptation
of past experiences in order to find solutions to current
problems.

@@

\subsection{CBR in Software Engineering}
Due to the advantages of CBR, numerous papers in SE
have applied this technique.    CBR  has been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}. 
Researchers have found that, in some domains, they are competitive with  closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. 
CBR is also useful when dealing with spare data sets~\cite{1438374} of with 
partial knowledge  at, for example,  an early stage of a project)~\cite{walkerden1999empirical}.

@@

Secondly, HOW   limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.

Turning now to the displacement {\em direction},
our experiments show that it is useful to  carefully reflect on exactly where HOW should  push each test case.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 20. In summary, 
If  $Z_j$ is an independent variable within the
test case $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z_j$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.

@@

    \caption{Training and test {\em data set properties} for the Jureczko data sets,
    sorted in ascending order of \% defective examples.
    On the right-hand-side, we show the {\em results from learning}.
    Data is ``good'' if it has   recalls over 60\% and false alarms under 40\%
(and note that, after tuning, there are more ``good'' than before).
Data   marked with ``$\star$'' show large improvements in performance, after tuning.
Data   marked with ``$\times$'' are not ``good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
}\label{fig:j}
\end{figure*}

@@

Turning now to the {\em test} data, this is passed to the { predictor}
to generate a baseline {\em before}  score.
If our { predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data {\em test'};
then (3)~return data on the {\em before,after} predictions.

@@

The clear advantage of the approach of \fig{work} is that it can be readily and automatically
applied to domains with historical data. That said, the obvious drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor predictors}, it is hence wise to demand certain standards for the {\em
satisfactory} threshold used in this \fig{work}.

For example, this study's data from two types of domains: the Jureczko   object-oriented data  
and the Seigmund configuration data. 
For the Seigmund data, the plans generated by HOW advise on how  to change the   configurations settings within a Makefile
(in order to make resulting compiled
code  run faster). Here, performance can be measured in terms of  difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 100*(1- \frac{abs(a - p)}{a})$.
This paper  explores    six Seigmund configuration data sets:  Berkeley DB (Java and Windows versions), Apache, SQLite, LLVM, and
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. 
  After splitting that data into equal sized train:test groups, a Random Forest
  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of
%issue 24
\[s=\{99.9, 99.8, 99.4, 99.1, 96,1\}\]
That is, we can be very confident that the predictors from the Seigmund data can assess
HOW's plans.
(Aside: if the reader doubts that such high scores are achievable, we note that these scores are 
consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)

@@

In summary, while we cannot trust predictors from some of our defect data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
One important detail to be stressed here is that, when we applied    SMOTE-ing and
parameter tunings, those techniques should be applied the training data and {\em not}
the test data;  
i.e.
we took care that no clues from the test set were ever used in this tuning process.

@@

\subsection{Data}
We applied HOW's plans to all  Seigmund data, and the five ``good''
  Jureczko   data sets {\em jedit, ivy, ant, lucene} and {\em poi}.
  The training and tests used for Jurecko were display in \fig{j}. For these data
  sets, training was managed by a Random Forest (tuned in the SMOTED training data using differential
  evolution). \fig{ck} shows Jureczko data attributes.

   \fig{cpm} shows the  Seigmund data attributes.
   There were more diverse than the 
    Jurecko data since they
   contain the results of applying many  mutators
  to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  Windows  versions),  Apache,  SQLite,  LLVM, and x264.

@@

   For both the Siegmind and Jureczko, we designed a procedure such we had at least
   20 samples from the \fig{work} procedure for  each data set (the value of 20 was choosen as the minimum use sample size via the central limit theorem):
   \bi
   \item
  For the Siegmind data, 
  25 times, 
  we randomized the order of the data, trained on one half (via
  Random Forest Regressor), then tested on the other.
  \item
  For the Jureczko, if we just  used the training and test sets of  \fig{j}, that would only
  give us sample per data set. Such a singleton sample could hide important details since the methods
  used for parameter tuning (SMOTE and differential evolution) use some stochastic sub-routines.
  Hence, for each  of the training sets of \fig{j}, we ran SMOTE+differential evolution 25 times to learn
  parameters for Random Forests.
  \ei
\fig{where} defined two control parameters for HOW:
\bi
\item  $F$   controls the size of the displacement;
\item   $B$   controls how many  attributes are displaced.
\ei
The   25 repeats defined above were executed for various values of $F,B$.
A Scott-Knott procedure was used to test for statistically significant differences in the results.
The Scott-Knott
procedure recommended by Mittas \& Angelis in their 2013
IEEE TSE paper~\cite{mittas13}.  This method
sorts a list of $l$ treatments with $ls$ measurements by their median
score. It then
splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
Scott-Knott  applies an  hypothesis test $H$ to check
if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.

@@

\begin{figure*}[!t]
\begin{center}
\begin{minipage}{.44\linewidth}
%\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Apache}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.79  &  0.06 & \quart{0}{11}{7}{49} \\
\hline  2 & F=0.75, B=50 &    0.82  &  0.07 & \quart{7}{14}{13}{49} \\
\hline  3 &  F=0.75, B=100  &    0.85  &  0.06 & \quart{11}{12}{19}{49} \\
  3 & F=0.5, B=25 &    0.87  &  0.03 & \quart{19}{6}{23}{49} \\
  3 & F=0.5, B=50 &    0.87  &  0.07 & \quart{17}{14}{23}{49} \\
\hline  4 &   F=0.5, B=100  &    0.91  &  0.07 & \quart{21}{14}{31}{49} \\
\hline  5 & F=0.25, B=25 &    0.92  &  0.05 & \quart{29}{10}{33}{49} \\
  5 & F=0.25, B=50 &    0.94  &  0.04 & \quart{35}{8}{37}{49} \\
  5 &  F=0.25,B=100   &    0.95  &  0.04 & \quart{35}{8}{39}{49} \\
\hline  6 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}

\vspace{2mm}\noindent {\scriptsize \textbf{BDBC}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=50 &    0.2  &  0.03 & \quart{0}{1}{1}{49} \\
  1 & F=0.75, B=25 &    0.2  &  0.05 & \quart{0}{3}{1}{49} \\
\hline  2 &  F=0.75, B=100  &    0.22  &  0.07 & \quart{1}{5}{2}{49} \\
\hline  3 & F=0.5, B=25 &    0.39  &  0.06 & \quart{11}{4}{12}{49} \\
  3 & F=0.5, B=50 &    0.4  &  0.05 & \quart{12}{3}{13}{49} \\
\hline  4 &   F=0.5,B=100   &    0.45  &  0.08 & \quart{14}{4}{16}{49} \\
\hline  5 & F=0.25, B=25 &    0.65  &  0.05 & \quart{27}{3}{28}{49} \\
  5 & F=0.25, B=50 &    0.66  &  0.05 & \quart{27}{3}{29}{49} \\
\hline  6 &  F=0.25,B=100   &    0.69  &  0.05 & \quart{29}{3}{31}{49} \\
\hline  7 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}

\vspace{2mm}\noindent{\scriptsize \textbf{BDBJ}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.64  &  0.1 & \quart{1}{12}{6}{49} \\
  1 & F=0.75, B=50 &    0.68  &  0.14 & \quart{0}{17}{10}{49} \\
\hline  2 &  F=0.75, B=100   &    0.74  &  0.1 & \quart{12}{12}{18}{49} \\
  2 & F=0.5, B=25 &    0.76  &  0.09 & \quart{15}{11}{20}{49} \\
  2 & F=0.5, B=50 &    0.78  &  0.17 & \quart{15}{21}{23}{49} \\
  2 &   F=0.5,B=100   &    0.8  &  0.07 & \quart{20}{9}{25}{49} \\
\hline  3 & F=0.25, B=25 &    0.85  &  0.05 & \quart{30}{6}{31}{49} \\
\hline  4 & F=0.25, B=50 &    0.89  &  0.09 & \quart{30}{11}{36}{49} \\
  4 &  F=0.25, B=100  &    0.9  &  0.06 & \quart{35}{7}{37}{49} \\
\hline  5 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}
%\noindent\begin{minipage}{0.5\textwidth}
%  \flushleft
\end{minipage}~~~~~\begin{minipage}{.44\linewidth}
\noindent
{\scriptsize \textbf{LLVM}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25 &    0.92  &  0.02 & \quart{0}{11}{5}{49} \\
  1 & F=0.75, B=50 &    0.93  &  0.02 & \quart{5}{11}{11}{49} \\
  1 &  F=0.75, B=100   &    0.93  &  0.02 & \quart{11}{11}{11}{49} \\
\hline  2 & F=0.5, B=25 &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 & F=0.5, B=50 &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 &   F=0.5, B=100   &    0.96  &  0.01 & \quart{22}{5}{27}{49} \\
\hline  3 & F=0.25, B=25 &    0.97  &  0.01 & \quart{33}{5}{33}{49} \\
  3 & F=0.25, B=50 &    0.98  &  0.0 & \quart{38}{0}{38}{49} \\
  3 &  F=0.25, B=100  &    0.98  &  0.01 & \quart{38}{6}{38}{49} \\
\hline  4 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}

>>>>

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, those improvements   reduce
 the expected values of defect counts and  runtimes to    
 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values, respectively.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, case-based reasoning.
  \end{IEEEkeywords}

@@

This is an open and pressing issue.
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself; e.g.
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
To generate those plans, we propose the ``HOW'' case-based reasoner (CBR).
HOW's  ``plans'' are  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every test case $Z$, HOW proposes a  plan $\delta$ to
  adjust   attributes $Z_j$:
\[
Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, HOW might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).
HOW's  plans displace the test case into some region
of the training data. This is important since we will use that training data
to build a predictor to assess the impact of that plan.
To put that another way, when a novel test case is presented,
HOW says 
\begin{quote}{\em ``I may not know too much about that, but once you change into something like
what I've already seen, then I can suggest how to improve it''.}
\end{quote}
After generating the plans, the next question is ``how to assess them?''.  
Some organizations have the resources to 
run repeated trials to assess  different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same for specification that was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
then check in 12 months time
 which teams have fewer defects than the others.

@@

This paper proposes another, less resource intensive,  method for assessing the effectiveness of proposed
changes to a project.  
We  use a prediction system (built by data mining) as a verification
tool for planners. If for certain domains data miners can build
predictors with adequate performance, then those predictors can assess the value
of proposed changes.
This paper applies this proposal to show that   HOW's plans are quite effective and  can
 significantly reduce the  predicted
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
The rest of this paper presents the background to this work in Section~2; the design of HOW in Section~3;
then our experimental results  in Section~4.  Our conclusion is that we can plan for the future in
software engineering, provided the domain data supports adequate predictors.

@@

However, not all domains come with ready-to-use models. For example, 
consider software defect prediction and all the intricate issues that may lead to defects in a product. A model that includes {\em all} those
potential issues would be very large and complex. Further,
the empirical data required to validate any/all parts
of that model can be hard to find.

@@

In summary, for domains with readily accessible models, we recommend
the kinds of tools that are widely used in the search-based
software engineering community such as GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like HOW.

\subsection{Planning and AI}
HOW is a {\em probabilistic} planner, which is a different approach to that taken by
classical AI. 
That kind of planning is a logical procedure~\cite{Fikes1971}
that seeks an ordering on {\em operators} to take some domain
{\em state} from a {\em start} state to a {\em  goal} state.
This classical logical approach is known to suffer from
computational bottlenecks~\cite{Bylander1994}. Hence,
like this paper,
many researchers take a probabilistic approach.
HOW's commitment to minimal displacement (and the fully-automated
modeling approach of this paper) makes it stand apart from, say,
  the Bayesian/Markov chain approach to planning by Dean et al.~\cite{dean1995planning}.

\subsection{Case-based Reasoning}
Case-based reasoning (CBR) has been extensively
explored in the AI literature
(where it has also been called
 reasoning via analogy or instance-based reasoning).
CBR seeks to emulate human recollection and adaptation
of past experiences in order to find solutions to current
problems.

@@

\subsection{CBR in Software Engineering}
Due to the advantages of CBR, numerous papers in SE
have applied this technique.    CBR  has been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}. 
Researchers have found that, in some domains, they are competitive with  closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. 
CBR is also useful when dealing with sparse data sets~\cite{1438374} or with 
partial knowledge  at, for example,  an early stage of a project)~\cite{walkerden1999empirical}.

@@

Secondly, HOW   limits displacement
to very small regions in the space of the training data (never more than the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.

Turning now to the displacement {\em direction},
our experiments show that it is useful to  carefully reflect on exactly where HOW should  push each test case.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 20. In summary, 
if  $Z_j$ is an independent variable within the
test case $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z_j$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.

@@

    \caption{Training and test {\em data set properties} for the Jureczko data sets,
    sorted in ascending order of \% defective examples.
    On the right-hand-side, we show the {\em results from learning}.
    Data is ``good'' if it has recall over 60\% and false alarm under 40\%
(and note that, after tuning, there are more ``good'' than before).
Data   marked with ``$\star$'' show large improvements in performance, after tuning.
Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
}\label{fig:j}
\end{figure*}

@@

Turning now to the {\em test} data, this is passed to the { predictor}
to generate a baseline {\em before}  score.
If our { predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data $test'$;
then (3)~return data on the {\em before,after} predictions.

@@

The clear advantage of the approach of \fig{work} is that it can be readily and automatically
applied to domains with historical data. That said, the obvious drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for the {\em
satisfactory} threshold used in this \fig{work}.

For example, this study's data arise from two types of domains: the Jureczko   object-oriented data  
and the Seigmund configuration data. 
For the Seigmund data, the plans generated by HOW advise on how  to change the   configurations settings within a Makefile
(in order to make resulting compiled
code  run faster). Here, performance can be measured in terms of  difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 100*(1- \frac{abs(a - p)}{a})$.
This paper  explores    six Seigmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. 
  After splitting that data into equal sized train:test groups, a Random Forest
  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of
%issue 24
\[s=\{99.9, 99.8, 99.4, 99.1, 96,1\}\]
That is, we can be very confident that the predictors from the Seigmund data can assess
HOW's plans.
(Aside: if the reader doubts that such high scores are achievable, we note that these scores are 
consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)

@@

In summary, while we cannot trust predictors from some of our defect data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
One important detail to be stressed here is that, when we applied    SMOTE-ing and
parameter tunings, those techniques should be applied to the training data and {\em not}
the test data;  
i.e.
we took care that no clues from the test set were ever used in this tuning process.

@@

\subsection{Data}
We applied HOW's plans to all  Seigmund data, and the five ``good''
  Jureczko   data sets {\em jedit, ivy, ant, lucene,} and {\em poi}.
  The training and tests used for Jurecko were display in \fig{j}. For these data
  sets, training was managed by a Random Forest (with tuning the SMOTED training data using differential
  evolution). \fig{ck} shows Jureczko data attributes.

   \fig{cpm} shows the  Seigmund data attributes.
   They were more diverse than the 
    Jurecko data since they
   contain the results of applying many  mutators
  to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  C versions),  Apache,  SQLite,  LLVM, and x264.

@@

   For both the Siegmind and Jureczko, we designed a procedure such that we had at least
   20 samples from the \fig{work} procedure for  each data set (the value of 20 was choosen as the minimum use sample size via the central limit theorem):
   \bi
   \item
  For the Siegmind data, 
  25 times, 
  we randomized the order of the data, trained on one half (via
  Random Forest Regressor), then tested on the other.
  \item
  For the Jureczko, if we just  used the training and test sets of  \fig{j}, that would only
  give us sample per data set. Such a singleton sample could hide important details since the methods
  used for parameter tuning (SMOTE and differential evolution) use some stochastic sub-routines.
  Hence, for each  of the training sets of \fig{j}, we ran SMOTE+differential evolution 25 times to learn
  parameters for Random Forests.
  \ei
\fig{where} defined two control parameters for HOW:
\bi
\item  $F$   controls the size of the displacement;
\item   $B$   controls how many  attributes are displaced.
\ei
The   25 repeats defined above were executed for various values of $F,B$.
A Scott-Knott procedure was used to test for statistically significant differences in the results.
The Scott-Knott
procedure recommended by Mittas \& Angelis in their 2013
IEEE TSE paper~\cite{mittas13}.  This method
sorts a list of $l$ treatments with $ls$ measurements by their median
score. It then
splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
Scott-Knott  applies an  hypothesis test $H$ to check
if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.

@@

\begin{figure*}[!t]
\begin{center}
\begin{minipage}{.44\linewidth}
%\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Apache}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25\% &    0.79  &  0.06 & \quart{0}{11}{7}{49} \\
\hline  2 & F=0.75, B=50\% &    0.82  &  0.07 & \quart{7}{14}{13}{49} \\
\hline  3 &  F=0.75, B=100\% &    0.85  &  0.06 & \quart{11}{12}{19}{49} \\
  3 & F=0.5, B=25\% &    0.87  &  0.03 & \quart{19}{6}{23}{49} \\
  3 & F=0.5, B=50\% &    0.87  &  0.07 & \quart{17}{14}{23}{49} \\
\hline  4 &   F=0.5, B=100\% &    0.91  &  0.07 & \quart{21}{14}{31}{49} \\
\hline  5 & F=0.25, B=25\% &    0.92  &  0.05 & \quart{29}{10}{33}{49} \\
  5 & F=0.25, B=50\% &    0.94  &  0.04 & \quart{35}{8}{37}{49} \\
  5 &  F=0.25,B=100\% &    0.95  &  0.04 & \quart{35}{8}{39}{49} \\
\hline  6 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}

\vspace{2mm}\noindent {\scriptsize \textbf{BDBC}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=50\% &    0.2  &  0.03 & \quart{0}{1}{1}{49} \\
  1 & F=0.75, B=25\% &    0.2  &  0.05 & \quart{0}{3}{1}{49} \\
\hline  2 &  F=0.75, B=100\% &    0.22  &  0.07 & \quart{1}{5}{2}{49} \\
\hline  3 & F=0.5, B=25\% &    0.39  &  0.06 & \quart{11}{4}{12}{49} \\
  3 & F=0.5, B=50\% &    0.4  &  0.05 & \quart{12}{3}{13}{49} \\
\hline  4 &   F=0.5,B=100\% &    0.45  &  0.08 & \quart{14}{4}{16}{49} \\
\hline  5 & F=0.25, B=25\% &    0.65  &  0.05 & \quart{27}{3}{28}{49} \\
  5 & F=0.25, B=50\% &    0.66  &  0.05 & \quart{27}{3}{29}{49} \\
\hline  6 &  F=0.25,B=100\% &    0.69  &  0.05 & \quart{29}{3}{31}{49} \\
\hline  7 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}

\vspace{2mm}\noindent{\scriptsize \textbf{BDBJ}\\[0.1cm]}
{\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25\% &    0.64  &  0.1 & \quart{1}{12}{6}{49} \\
  1 & F=0.75, B=50\% &    0.68  &  0.14 & \quart{0}{17}{10}{49} \\
\hline  2 &  F=0.75, B=100\% &    0.74  &  0.1 & \quart{12}{12}{18}{49} \\
  2 & F=0.5, B=25\% &    0.76  &  0.09 & \quart{15}{11}{20}{49} \\
  2 & F=0.5, B=50\% &    0.78  &  0.17 & \quart{15}{21}{23}{49} \\
  2 &   F=0.5,B=100\% &    0.8  &  0.07 & \quart{20}{9}{25}{49} \\
\hline  3 & F=0.25, B=25\% &    0.85  &  0.05 & \quart{30}{6}{31}{49} \\
\hline  4 & F=0.25, B=50\% &    0.89  &  0.09 & \quart{30}{11}{36}{49} \\
  4 &  F=0.25, B=100\% &    0.9  &  0.06 & \quart{35}{7}{37}{49} \\
\hline  5 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}
%\noindent\begin{minipage}{0.5\textwidth}
%  \flushleft
\end{minipage}~~~~~\begin{minipage}{.44\linewidth}
\noindent
{\scriptsize \textbf{LLVM}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{llrrc}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & F=0.75, B=25\% &    0.92  &  0.02 & \quart{0}{11}{5}{49} \\
  1 & F=0.75, B=50\% &    0.93  &  0.02 & \quart{5}{11}{11}{49} \\
  1 &  F=0.75, B=100\% &    0.93  &  0.02 & \quart{11}{11}{11}{49} \\
\hline  2 & F=0.5, B=25 &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 & F=0.5, B=50\% &    0.95  &  0.01 & \quart{16}{6}{22}{49} \\
  2 &   F=0.5, B=100\% &    0.96  &  0.01 & \quart{22}{5}{27}{49} \\
\hline  3 & F=0.25, B=25\% &    0.97  &  0.01 & \quart{33}{5}{33}{49} \\
  3 & F=0.25, B=50\% &    0.98  &  0.0 & \quart{38}{0}{38}{49} \\
  3 &  F=0.25, B=100\% &    0.98  &  0.01 & \quart{38}{6}{38}{49} \\
\hline  4 &     Baseline &    1.0  &  0.0 & \quart{49}{0}{49}{49} \\
\hline \end{tabular}}
%\end{minipage}
